---
title: "Transit Cost Project"
author: "Arthur Gailes"
date: "1/7/2021"
output: github_document
editor_options: 
  chunk_output_type: inline
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F)
sapply(c('tidyverse','tidytuesdayR','r2d3','here'), require, character.only=T)
knitr::opts_knit$set(root.dir = here())
```

## Transit Cost Project

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r download}
tt <- tt_load("2021-01-05")
trans_raw <- tt[[1]]
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
# look at data and save to dataframe
tt %>% 
  map(glimpse)

```

# Wrangle data

Calculate average timeline for projects by start year, for projects that have already started.
```{r}
trans <- trans_raw %>% 
  mutate(start_year = as.numeric(year), end_year=as.numeric(end_year), 
    year_length = end_year-start_year) %>% 
  filter(!is.na(year_length), start_year <2021) %>% 
  group_by(start_year) %>% 
  summarize(year_length = mean(year_length))
trans
```

# R2D3

Visualize project timeline by year - simple.

```{d3 data=trans, fig.show='hold'}

const fig = {
  margin: {top:50,right:20, bottom:20, left: 40}
};
fig.height = height - fig.margin.top - fig.margin.bottom;
fig.width = width - fig.margin.right - fig.margin.left;

var x = d3.scaleLinear()
          .range([0, fig.width])
          .domain(d3.extent(data, d=> d.start_year))
          ;
var y = d3.scaleLinear()
          .range([fig.height, fig.margin.top])
          .domain([0, d3.max(data, d=> d.year_length)])
          ;
const xAxis = d3.axisBottom(x)
  .tickFormat(d3.format(".0f"))
  .ticks(20)
  ;
          
// Add bars
svg
  .attr('height', height)
  .attr('width', width)
  .append('g')
  .attr('transform','translate(' + fig.margin.left + ',' + 0 + ')')
  .selectAll('rect')
  .data(data)
  .join(enter=> enter
    .append('rect')
      .attr('width', fig.width/(d3.max(data, d1=> d1.start_year) - d3.min(data, d1=>d1.start_year)))
      .attr('height', d=> fig.height - y(d.year_length))
      .attr('y', d=> y(d.year_length))
      .attr('x',d=> x(d.start_year))
      .attr('fill', 'red')
      .attr('stroke','black')
      )
      
// Axes
svg.append('g')
  .attr('transform','translate(' + fig.margin.left + ',' + fig.height + ')')
  .call(xAxis)
  .selectAll('text')
  .style('text-anchor','end')
  .attr('transform','rotate(-65)')
  .attr("dx", "-.8em")
  .attr("dy", ".15em")
  ;
  
svg.append('g')
  .attr('transform','translate(' + fig.margin.left + ')')
  .call(d3.axisLeft(y))
  
svg.append('text')
  .attr('class','axisLabel')
  .attr('x',fig.width/2 - fig.margin.right)
  .attr('y', height)
  .text('Project Start Year')
;
svg.append('text')
  .attr('class','axisLabel')
  .attr('x',0-height/1.5)
  .attr('y', 12)
  .attr('transform','rotate(-90)')
  .text('Years to Complete')
;
//title
svg.append('text')
  .attr('class','axisLabel')
  .attr('x',150)
  .attr('y', fig.margin.bottom)
  .text('Number of years for Transit Projects to complete')
;
  
```

